Housing Market Institutions Drive Race and Ethnicity Differences in Energy Consumption

### 2015-04-07 ConferencePaper/DraftEdwardsWoods.tex
Added descriptive text to the behavior differences section.

Table 
 \ref{tab:RaceVStruct}
shows the count of observations by Race/Ethnicty and structure after the basic data handling described above.  Please note the dominance of single family detached homes, shown as SFDetached in the table and non-Hispanic Caucasians, shown as FALSEWt in the table.  These counts are unweighted and do not take into account the stratefied sampling scheme used to collect the data.

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Sat Apr  4 20:18:51 2015
\begin{table}[ht]
\centering
\caption{Count of Observations by Race and Structure Type} 
\label{tab:RaceVStruct}
{\small
\begin{tabular}{rrrrrr}
  \hline
 & Mobile & SFDetached & SFAttached & SmApartment & LgApartment \\ 
  \hline
FALSEWt & 371 & 5694 & 544 & 430 & 951 \\ 
  FALSEAfAm &  24 & 754 & 121 & 174 & 382 \\ 
  FALSEAsian &   1 & 231 &  48 &  52 & 113 \\ 
  FALSEMulti &   6 &  87 &  14 &  12 &  27 \\ 
  FALSENativeAm &   7 &  35 &   9 &   7 &  15 \\ 
  FALSEOther &   2 &  59 &   8 &  13 &  29 \\ 
  FALSEPacific &   1 &  22 &   1 &   2 &  10 \\ 
  TRUEAfAm &   3 &   8 &  10 &   5 &   8 \\ 
  TRUEAsian &   0 &   5 &   0 &   0 &   1 \\ 
  TRUEMulti &   1 &  10 &   2 &   4 &   4 \\ 
  TRUENativeAm &   2 &  17 &   3 &   6 &   9 \\ 
  TRUEOther &   6 &  41 &  12 &  16 &  24 \\ 
  TRUEPacific &   0 &   2 &   1 &   1 &   0 \\ 
  TRUEWt &  97 & 730 & 105 & 184 & 332 \\ 
   \hline
\end{tabular}
}
\end{table}
Many of the structure type/ race and ethnicity combinations are relatively rare, e.g., Hispanic Asians living in mobile homes.  We have not collapsed any of these  categories.  Future research will look at collapsing some of the rare cells so that there are a more managable number of both structure types and race/ethnicity combinations.

Table
\ref{tab:RegionVStruct}
gives a further description of the data is distributed around the country.  Please note that both the Western and North Eastern area have a larger fraction of large apartment buldings, LgApartment in the table, than the other two regions.

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Sat Apr  4 20:18:51 2015
\begin{table}[ht]
\centering
\caption{Count of Observations by Region and Structure Type} 
\label{tab:RegionVStruct}
{\small
\begin{tabular}{rrrrr}
  \hline
 & NE & MidWest & South & West \\ 
  \hline
Mobile &  47 &  92 & 245 & 137 \\ 
  SFDetached & 1173 & 2080 & 2685 & 1757 \\ 
  SFAttached & 218 & 162 & 257 & 241 \\ 
  SmApartment & 326 & 145 & 226 & 209 \\ 
  LgApartment & 470 & 323 & 603 & 509 \\ 
   \hline
\end{tabular}
}
\end{table}
The table shows the range of climate zones and houseing law covered by the survey.  It also shows the reasonablness of our hypotheses that location and structure type choices are endogenous.

@@

\begin{figure}
\begin{center}
\caption{Daytime Temp When Away and Race/Ethnicity (Winter)}
\label{fig:DayAwayRace}
\includegraphics{DraftEdwardsWoods-013}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Nighttime Temp and Race/Ethnicity (Winter)}
\label{fig:NightRace}
\includegraphics{DraftEdwardsWoods-014}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Daytime Temp Home by Race/Ethnicity (Summer)}
\label{fig:HomeRaceS}
\includegraphics{DraftEdwardsWoods-015}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Day Temp Away by Race/Ethnicity (Summer)}
\label{fig:AwayRaceS}
\includegraphics{DraftEdwardsWoods-016}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Nighttime Temp and Race/Ethnicity (Summer)}
\label{fig:NightRaceS}
\includegraphics{DraftEdwardsWoods-017}
\end{center}
\end{figure}

@@

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Sat Apr  4 20:20:08 2015
\begin{table}[ht]
\centering
\caption{HUD Complaints per 100,000 Population} 
\label{tab:HUDComplaints}
{\small
\begin{tabular}{llr}
  \hline
State & Population & Reports (per 100,000) \\ 
  \hline
Alabama & 4,677,464 & 7.50 \\ 
  Alaska & 688,125 & 13.01 \\ 
  Arizona & 6,499,377 & 3.68 \\ 
  Arkansas & 2,867,764 & 9.26 \\ 
  California & 36,580,371 & 3.03 \\ 
  Colorado & 4,935,213 & 1.97 \\ 
  Connecticut & 3,502,932 & 20.50 \\ 
  Delaware & 876,211 & 14.00 \\ 
  District of Columbia & 590,074 & 14.00 \\ 
  Florida & 18,423,878 & 3.93 \\ 
  Georgia & 9,697,838 & 2.06 \\ 
  Hawaii & 1,287,481 & 13.01 \\ 
  Idaho & 1,527,506 & 9.94 \\ 
  Illinois & 12,842,954 & 2.88 \\ 
  Indiana & 6,388,309 & 9.14 \\ 
  Iowa & 2,993,987 & 19.92 \\ 
  Kansas & 2,797,375 & 10.31 \\ 
  Kentucky & 4,287,931 & 7.50 \\ 
  Louisiana & 4,451,513 & 9.26 \\ 
  Maine & 1,319,691 & 20.50 \\ 
  Maryland & 5,658,655 & 14.00 \\ 
  Massachusetts & 6,543,595 & 4.51 \\ 
  Michigan & 10,002,486 & 4.97 \\ 
  Minnesota & 5,230,567 & 19.92 \\ 
  Mississippi & 2,940,212 & 7.50 \\ 
  Missouri & 5,956,335 & 4.99 \\ 
  Montana & 968,035 & 9.94 \\ 
  Nebraska & 1,781,949 & 10.31 \\ 
  Nevada & 2,615,772 & 5.08 \\ 
  New Hampshire & 1,321,872 & 20.50 \\ 
  New Jersey & 8,663,398 & 2.34 \\ 
  New Mexico & 1,986,763 & 5.08 \\ 
  New York & 19,467,789 & 4.60 \\ 
  North Carolina & 9,247,134 & 4.22 \\ 
  North Dakota & 641,421 & 19.92 \\ 
  Ohio & 11,528,072 & 9.14 \\ 
  Oklahoma & 3,644,025 & 9.26 \\ 
  Oregon & 3,782,991 & 13.01 \\ 
  Pennsylvania & 12,566,368 & 1.87 \\ 
  Rhode Island & 1,053,502 & 20.50 \\ 
  South Carolina & 4,503,280 & 4.22 \\ 
  South Dakota & 804,532 & 19.92 \\ 
  Tennessee & 6,240,456 & 2.61 \\ 
  Texas & 24,304,290 & 4.18 \\ 
  Utah & 2,727,343 & 9.94 \\ 
  Vermont & 621,049 & 20.50 \\ 
  Virginia & 7,795,424 & 2.03 \\ 
  Washington & 6,566,073 & 13.01 \\ 
  West Virginia & 1,814,873 & 14.00 \\ 
  Wisconsin & 5,627,610 & 1.81 \\ 
  Wyoming & 532,981 & 9.94 \\ 
   \hline
\end{tabular}
}
\end{table}
\section{Conditional Demand Estimation}

@@

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Sat Apr  4 20:20:09 2015
{\small
\begin{longtable}{rrrrr}
\caption{Orthodox kWh Model} \\ 
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
Income & 0.0086 & 0.0016 & 5.3662 & 0.0000 \\ 
  Income:EthRaceFALSEAfAm & 0.0031 & 0.0027 & 1.1389 & 0.2548 \\ 
  Income:EthRaceFALSEAsian & -0.0243 & 0.0034 & -7.1646 & 0.0000 \\ 
  Income:EthRaceFALSEMulti & -0.0168 & 0.0069 & -2.4259 & 0.0153 \\ 
  Income:EthRaceFALSENativeAm & 0.0117 & 0.0123 & 0.9493 & 0.3425 \\ 
  Income:EthRaceFALSEOther & -0.0249 & 0.0083 & -2.9909 & 0.0028 \\ 
  Income:EthRaceFALSEPacific & -0.0163 & 0.0138 & -1.1807 & 0.2378 \\ 
  Income:EthRaceTRUEAfAm & -0.0120 & 0.0165 & -0.7271 & 0.4672 \\ 
  Income:EthRaceTRUEAsian & -0.0365 & 0.0454 & -0.8032 & 0.4219 \\ 
  Income:EthRaceTRUEMulti & -0.0291 & 0.0212 & -1.3745 & 0.1693 \\ 
  Income:EthRaceTRUENativeAm & -0.0187 & 0.0148 & -1.2599 & 0.2077 \\ 
  Income:EthRaceTRUEOther & -0.0262 & 0.0099 & -2.6335 & 0.0085 \\ 
  Income:EthRaceTRUEPacific & -0.0122 & 0.0380 & -0.3221 & 0.7474 \\ 
  Income:EthRaceTRUEWt & -0.0192 & 0.0031 & -6.2907 & 0.0000 \\ 
  StrTenureOwnSFDetached:TOTSQFT\_EN:HDD65 & 0.0001 & 0.0000 & 4.1043 & 0.0000 \\ 
  StrTenureOwnLgApartment:TOTSQFT\_EN:HDD65 & -0.0001 & 0.0001 & -0.9580 & 0.3381 \\ 
  StrTenureOwnMobile:TOTSQFT\_EN:HDD65 & 0.0004 & 0.0001 & 5.9838 & 0.0000 \\ 
  StrTenureOwnSFAttached:TOTSQFT\_EN:HDD65 & -0.0000 & 0.0000 & -0.5516 & 0.5812 \\ 
  StrTenureOwnSmApartment:TOTSQFT\_EN:HDD65 & 0.0000 & 0.0001 & 0.5438 & 0.5866 \\ 
  StrTenureRentLgApartment:TOTSQFT\_EN:HDD65 & -0.0002 & 0.0001 & -4.0031 & 0.0001 \\ 
  StrTenureRentMobile:TOTSQFT\_EN:HDD65 & 0.0004 & 0.0002 & 1.7633 & 0.0779 \\ 
  StrTenureRentSFAttached:TOTSQFT\_EN:HDD65 & 0.0000 & 0.0001 & 0.6010 & 0.5479 \\ 
  StrTenureRentSFDetached:TOTSQFT\_EN:HDD65 & 0.0000 & 0.0000 & 0.5515 & 0.5813 \\ 
  StrTenureRentSmApartment:TOTSQFT\_EN:HDD65 & -0.0001 & 0.0001 & -1.8580 & 0.0632 \\ 
  StrTenureOwnSFDetached:TOTSQFT\_EN:CDD65 & 0.0008 & 0.0000 & 19.2068 & 0.0000 \\ 
  StrTenureOwnLgApartment:TOTSQFT\_EN:CDD65 & -0.0002 & 0.0002 & -0.8168 & 0.4140 \\ 
  StrTenureOwnMobile:TOTSQFT\_EN:CDD65 & 0.0009 & 0.0001 & 7.3167 & 0.0000 \\ 
  StrTenureOwnSFAttached:TOTSQFT\_EN:CDD65 & 0.0006 & 0.0001 & 5.0127 & 0.0000 \\ 
  StrTenureOwnSmApartment:TOTSQFT\_EN:CDD65 & 0.0001 & 0.0003 & 0.4022 & 0.6875 \\ 
  StrTenureRentLgApartment:TOTSQFT\_EN:CDD65 & 0.0000 & 0.0001 & 0.3926 & 0.6946 \\ 
  StrTenureRentMobile:TOTSQFT\_EN:CDD65 & 0.0024 & 0.0004 & 6.3205 & 0.0000 \\ 
  StrTenureRentSFAttached:TOTSQFT\_EN:CDD65 & 0.0007 & 0.0002 & 4.0210 & 0.0001 \\ 
  StrTenureRentSFDetached:TOTSQFT\_EN:CDD65 & 0.0008 & 0.0001 & 9.2680 & 0.0000 \\ 
  StrTenureRentSmApartment:TOTSQFT\_EN:CDD65 & 0.0005 & 0.0002 & 2.8365 & 0.0046 \\ 
   \hline
\hline
\label{tab:OrthoKWHLimited}
\end{longtable}
}
The top parameter, Income, shows the estimate for annual kWh per dollar of income for a household headed by a non-Hispanic Caucasian.  The remaining income related variables show the deviations from this case for Hispanic headed households, labeled TRUE in the table, and by the other self reported races.  Note that in all cases, with the exception of non-Hispanic Native Americans, the parameter estimates show less electricity used than non-Hispanic Caucasian households.  Only a few of these differences are statistically significant, non-Hispanic Asian and multi-ethnic households, as well as Hispanic Caucasian households and households that reported some other race.

@@

The full set of parameter estimates for the electricity model can be found in the appendix.  We will focus on a subset of the parameter estimates, focusing on the effects of race and ethnicity separatly and the effects of the other endogenous variables, structre and tenure and square footage.

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Sat Apr  4 20:20:11 2015
{\small
\begin{longtable}{rrrrl}
\caption{System Estimation of kWh Model: Race and Ethnicity} \\ 
  \hline
 & Estimate & Std. Error & t value & Sig \\ 
  \hline
Income:EthRaceFALSEAfAm & 0.0052 & 0.0029 & 1.7791 &   \\ 
  Income:EthRaceFALSEAsian & -0.0230 & 0.0035 & -6.4831 & *** \\ 
  Income:EthRaceFALSEMulti & -0.0157 & 0.0057 & -2.7584 & * \\ 
  Income:EthRaceFALSENativeAm & 0.0092 & 0.0211 & 0.4382 &   \\ 
  Income:EthRaceFALSEOther & -0.0237 & 0.0082 & -2.9038 & * \\ 
  Income:EthRaceFALSEPacific & -0.0129 & 0.0230 & -0.5625 &   \\ 
  Income:EthRaceTRUEAfAm & -0.0173 & 0.0139 & -1.2454 &   \\ 
  Income:EthRaceTRUEAsian & -0.0375 & 0.0101 & -3.7155 & ** \\ 
  Income:EthRaceTRUEMulti & -0.0331 & 0.0192 & -1.7208 &   \\ 
  Income:EthRaceTRUENativeAm & -0.0163 & 0.0107 & -1.5276 &   \\ 
  Income:EthRaceTRUEOther & -0.0274 & 0.0080 & -3.4318 & ** \\ 
  Income:EthRaceTRUEPacific & -0.0240 & 0.0659 & -0.3647 &   \\ 
  Income:EthRaceTRUEWt & -0.0185 & 0.0033 & -5.5782 & *** \\ 
   \hline
\hline
\label{tab:SystemKWHEthnicity}
\end{longtable}
}
Table 
% \ref{tab:SystemKWHEthnicity}
shows similarites with the results from the orthodox electrity model.  Non-Hispanic Asians, and Hispanic Caucasians and Hispanic heads of households that identify as some other race all show lower electricity use per dollar of income.   This is consistent in sign with the parameter estimates in table 
% \ref{tab:ParOrthKWH}
, but the statistically significant deviations from non-Hispanic caucasions for non Hispanic multi-ethnic heads of household has vanished.  The scale of the significant paremeter estimates are also roughly the same magnitude but attenuated towards zero.

>>>>

Table 
 \ref{tab:RaceVStruct}
shows the count of observations by Race/Ethnicity and structure after the basic data handling described above.  Please note the dominance of single family detached homes, shown as SFDetached in the table and non-Hispanic Caucasians, shown as FALSEWt in the table.  These counts are unweighted and do not take into account the stratified sampling scheme used to collect the data.

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Tue Apr  7 17:03:11 2015
\begin{table}[ht]
\centering
\caption{Count of Observations by Race and Structure Type} 
\label{tab:RaceVStruct}
{\small
\begin{tabular}{rrrrrr}
  \hline
 & Mobile & SFDetached & SFAttached & SmApartment & LgApartment \\ 
  \hline
FALSEWt & 371 & 5694 & 544 & 430 & 951 \\ 
  FALSEAfAm &  24 & 754 & 121 & 174 & 382 \\ 
  FALSEAsian &   1 & 231 &  48 &  52 & 113 \\ 
  FALSEMulti &   6 &  87 &  14 &  12 &  27 \\ 
  FALSENativeAm &   7 &  35 &   9 &   7 &  15 \\ 
  FALSEOther &   2 &  59 &   8 &  13 &  29 \\ 
  FALSEPacific &   1 &  22 &   1 &   2 &  10 \\ 
  TRUEAfAm &   3 &   8 &  10 &   5 &   8 \\ 
  TRUEAsian &   0 &   5 &   0 &   0 &   1 \\ 
  TRUEMulti &   1 &  10 &   2 &   4 &   4 \\ 
  TRUENativeAm &   2 &  17 &   3 &   6 &   9 \\ 
  TRUEOther &   6 &  41 &  12 &  16 &  24 \\ 
  TRUEPacific &   0 &   2 &   1 &   1 &   0 \\ 
  TRUEWt &  97 & 730 & 105 & 184 & 332 \\ 
   \hline
\end{tabular}
}
\end{table}
Many of the structure type/ race and ethnicity combinations are relatively rare, e.g., Hispanic Asians living in mobile homes.  We have not collapsed any of these  categories.  Future research will look at collapsing some of the rare cells so that there are a more manageable number of both structure types and race/ethnicity combinations.

Table
\ref{tab:RegionVStruct}
gives a further description of the data is distributed around the country.  Please note that both the Western and North Eastern area have a larger fraction of large apartment buildings, LgApartment in the table, than the other two regions.

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Tue Apr  7 17:03:11 2015
\begin{table}[ht]
\centering
\caption{Count of Observations by Region and Structure Type} 
\label{tab:RegionVStruct}
{\small
\begin{tabular}{rrrrr}
  \hline
 & NE & MidWest & South & West \\ 
  \hline
Mobile &  47 &  92 & 245 & 137 \\ 
  SFDetached & 1173 & 2080 & 2685 & 1757 \\ 
  SFAttached & 218 & 162 & 257 & 241 \\ 
  SmApartment & 326 & 145 & 226 & 209 \\ 
  LgApartment & 470 & 323 & 603 & 509 \\ 
   \hline
\end{tabular}
}
\end{table}
The table shows the range of climate zones and housing law covered by the survey.  It also shows the reasonableness of our hypotheses that location and structure type choices are endogenous.

@@

The figure clearly shows that non-Hispanic African American households report a higher daytime winter thermostat setting than the other non-Hispanic groups.  This could be locational.  This could be cultural, but it is a behavioral difference that could make African American housholds more energy intensive in this end-use.

The effects of this higher thermostat setting may be offset by less electricity use for plug loads or hot water service or by having smaller resdences, but it represents a different preference.

This difference in winter thermostat behavior follows through in summer thermostat preferences.  Figure \ref{fig:HomeRaceS} shows a summary for similar summer thermostat settings.  Note that non-Hispanic African Americans, Asian Americans and Native Americans all report lower summer thermostat settings than non-Hispanic and Hispanic Caucasian households.

\begin{figure}
\begin{center}
\caption{Daytime Temp Home by Race/Ethnicity (Summer)}
\label{fig:HomeRaceS}
\includegraphics{DraftEdwardsWoods-013}
\end{center}
\end{figure}

@@

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Tue Apr  7 17:03:48 2015
\begin{table}[ht]
\centering
\caption{HUD Complaints per 100,000 Population} 
\label{tab:HUDComplaints}
{\small
\begin{tabular}{llr}
  \hline
State & Population & Reports (per 100,000) \\ 
  \hline
Alabama & 4,677,464 & 7.50 \\ 
  Alaska & 688,125 & 13.01 \\ 
  Arizona & 6,499,377 & 3.68 \\ 
  Arkansas & 2,867,764 & 9.26 \\ 
  California & 36,580,371 & 3.03 \\ 
  Colorado & 4,935,213 & 1.97 \\ 
  Connecticut & 3,502,932 & 20.50 \\ 
  Delaware & 876,211 & 14.00 \\ 
  District of Columbia & 590,074 & 14.00 \\ 
  Florida & 18,423,878 & 3.93 \\ 
  Georgia & 9,697,838 & 2.06 \\ 
  Hawaii & 1,287,481 & 13.01 \\ 
  Idaho & 1,527,506 & 9.94 \\ 
  Illinois & 12,842,954 & 2.88 \\ 
  Indiana & 6,388,309 & 9.14 \\ 
  Iowa & 2,993,987 & 19.92 \\ 
  Kansas & 2,797,375 & 10.31 \\ 
  Kentucky & 4,287,931 & 7.50 \\ 
  Louisiana & 4,451,513 & 9.26 \\ 
  Maine & 1,319,691 & 20.50 \\ 
  Maryland & 5,658,655 & 14.00 \\ 
  Massachusetts & 6,543,595 & 4.51 \\ 
  Michigan & 10,002,486 & 4.97 \\ 
  Minnesota & 5,230,567 & 19.92 \\ 
  Mississippi & 2,940,212 & 7.50 \\ 
  Missouri & 5,956,335 & 4.99 \\ 
  Montana & 968,035 & 9.94 \\ 
  Nebraska & 1,781,949 & 10.31 \\ 
  Nevada & 2,615,772 & 5.08 \\ 
  New Hampshire & 1,321,872 & 20.50 \\ 
  New Jersey & 8,663,398 & 2.34 \\ 
  New Mexico & 1,986,763 & 5.08 \\ 
  New York & 19,467,789 & 4.60 \\ 
  North Carolina & 9,247,134 & 4.22 \\ 
  North Dakota & 641,421 & 19.92 \\ 
  Ohio & 11,528,072 & 9.14 \\ 
  Oklahoma & 3,644,025 & 9.26 \\ 
  Oregon & 3,782,991 & 13.01 \\ 
  Pennsylvania & 12,566,368 & 1.87 \\ 
  Rhode Island & 1,053,502 & 20.50 \\ 
  South Carolina & 4,503,280 & 4.22 \\ 
  South Dakota & 804,532 & 19.92 \\ 
  Tennessee & 6,240,456 & 2.61 \\ 
  Texas & 24,304,290 & 4.18 \\ 
  Utah & 2,727,343 & 9.94 \\ 
  Vermont & 621,049 & 20.50 \\ 
  Virginia & 7,795,424 & 2.03 \\ 
  Washington & 6,566,073 & 13.01 \\ 
  West Virginia & 1,814,873 & 14.00 \\ 
  Wisconsin & 5,627,610 & 1.81 \\ 
  Wyoming & 532,981 & 9.94 \\ 
   \hline
\end{tabular}
}
\end{table}
\section{Conditional Demand Estimation}

@@

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Tue Apr  7 17:03:48 2015
{\small
\begin{longtable}{rrrrr}
\caption{Orthodox kWh Model} \\ 
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
Income & 0.0086 & 0.0016 & 5.3662 & 0.0000 \\ 
  Income:EthRaceFALSEAfAm & 0.0031 & 0.0027 & 1.1389 & 0.2548 \\ 
  Income:EthRaceFALSEAsian & -0.0243 & 0.0034 & -7.1646 & 0.0000 \\ 
  Income:EthRaceFALSEMulti & -0.0168 & 0.0069 & -2.4259 & 0.0153 \\ 
  Income:EthRaceFALSENativeAm & 0.0117 & 0.0123 & 0.9493 & 0.3425 \\ 
  Income:EthRaceFALSEOther & -0.0249 & 0.0083 & -2.9909 & 0.0028 \\ 
  Income:EthRaceFALSEPacific & -0.0163 & 0.0138 & -1.1807 & 0.2378 \\ 
  Income:EthRaceTRUEAfAm & -0.0120 & 0.0165 & -0.7271 & 0.4672 \\ 
  Income:EthRaceTRUEAsian & -0.0365 & 0.0454 & -0.8032 & 0.4219 \\ 
  Income:EthRaceTRUEMulti & -0.0291 & 0.0212 & -1.3745 & 0.1693 \\ 
  Income:EthRaceTRUENativeAm & -0.0187 & 0.0148 & -1.2599 & 0.2077 \\ 
  Income:EthRaceTRUEOther & -0.0262 & 0.0099 & -2.6335 & 0.0085 \\ 
  Income:EthRaceTRUEPacific & -0.0122 & 0.0380 & -0.3221 & 0.7474 \\ 
  Income:EthRaceTRUEWt & -0.0192 & 0.0031 & -6.2907 & 0.0000 \\ 
  StrTenureOwnSFDetached:TOTSQFT\_EN:HDD65 & 0.0001 & 0.0000 & 4.1043 & 0.0000 \\ 
  StrTenureOwnLgApartment:TOTSQFT\_EN:HDD65 & -0.0001 & 0.0001 & -0.9580 & 0.3381 \\ 
  StrTenureOwnMobile:TOTSQFT\_EN:HDD65 & 0.0004 & 0.0001 & 5.9838 & 0.0000 \\ 
  StrTenureOwnSFAttached:TOTSQFT\_EN:HDD65 & -0.0000 & 0.0000 & -0.5516 & 0.5812 \\ 
  StrTenureOwnSmApartment:TOTSQFT\_EN:HDD65 & 0.0000 & 0.0001 & 0.5438 & 0.5866 \\ 
  StrTenureRentLgApartment:TOTSQFT\_EN:HDD65 & -0.0002 & 0.0001 & -4.0031 & 0.0001 \\ 
  StrTenureRentMobile:TOTSQFT\_EN:HDD65 & 0.0004 & 0.0002 & 1.7633 & 0.0779 \\ 
  StrTenureRentSFAttached:TOTSQFT\_EN:HDD65 & 0.0000 & 0.0001 & 0.6010 & 0.5479 \\ 
  StrTenureRentSFDetached:TOTSQFT\_EN:HDD65 & 0.0000 & 0.0000 & 0.5515 & 0.5813 \\ 
  StrTenureRentSmApartment:TOTSQFT\_EN:HDD65 & -0.0001 & 0.0001 & -1.8580 & 0.0632 \\ 
  StrTenureOwnSFDetached:TOTSQFT\_EN:CDD65 & 0.0008 & 0.0000 & 19.2068 & 0.0000 \\ 
  StrTenureOwnLgApartment:TOTSQFT\_EN:CDD65 & -0.0002 & 0.0002 & -0.8168 & 0.4140 \\ 
  StrTenureOwnMobile:TOTSQFT\_EN:CDD65 & 0.0009 & 0.0001 & 7.3167 & 0.0000 \\ 
  StrTenureOwnSFAttached:TOTSQFT\_EN:CDD65 & 0.0006 & 0.0001 & 5.0127 & 0.0000 \\ 
  StrTenureOwnSmApartment:TOTSQFT\_EN:CDD65 & 0.0001 & 0.0003 & 0.4022 & 0.6875 \\ 
  StrTenureRentLgApartment:TOTSQFT\_EN:CDD65 & 0.0000 & 0.0001 & 0.3926 & 0.6946 \\ 
  StrTenureRentMobile:TOTSQFT\_EN:CDD65 & 0.0024 & 0.0004 & 6.3205 & 0.0000 \\ 
  StrTenureRentSFAttached:TOTSQFT\_EN:CDD65 & 0.0007 & 0.0002 & 4.0210 & 0.0001 \\ 
  StrTenureRentSFDetached:TOTSQFT\_EN:CDD65 & 0.0008 & 0.0001 & 9.2680 & 0.0000 \\ 
  StrTenureRentSmApartment:TOTSQFT\_EN:CDD65 & 0.0005 & 0.0002 & 2.8365 & 0.0046 \\ 
   \hline
\hline
\label{tab:OrthoKWHLimited}
\end{longtable}
}
The top parameter, Income, shows the estimate for annual kWh per dollar of income for a household headed by a non-Hispanic Caucasian.  The remaining income related variables show the deviations from this case for Hispanic headed households, labeled TRUE in the table, and by the other self reported races.  Note that in all cases, with the exception of non-Hispanic Native Americans, the parameter estimates show less electricity used than non-Hispanic Caucasian households.  Only a few of these differences are statistically significant, non-Hispanic Asian and multi-ethnic households, as well as Hispanic Caucasian households and households that reported some other race.

@@

The full set of parameter estimates for the electricity model can be found in the appendix.  We will focus on a subset of the parameter estimates, focusing on the effects of race and ethnicity separately and the effects of the other endogenous variables, structure and tenure and square footage.

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Tue Apr  7 17:03:49 2015
{\small
\begin{longtable}{rrrrl}
\caption{System Estimation of kWh Model: Race and Ethnicity} \\ 
  \hline
 & Estimate & Std. Error & t value & Sig \\ 
  \hline
Income:EthRaceFALSEAfAm & 0.0052 & 0.0029 & 1.7791 &   \\ 
  Income:EthRaceFALSEAsian & -0.0230 & 0.0035 & -6.4831 & *** \\ 
  Income:EthRaceFALSEMulti & -0.0157 & 0.0057 & -2.7584 & * \\ 
  Income:EthRaceFALSENativeAm & 0.0092 & 0.0211 & 0.4382 &   \\ 
  Income:EthRaceFALSEOther & -0.0237 & 0.0082 & -2.9038 & * \\ 
  Income:EthRaceFALSEPacific & -0.0129 & 0.0230 & -0.5625 &   \\ 
  Income:EthRaceTRUEAfAm & -0.0173 & 0.0139 & -1.2454 &   \\ 
  Income:EthRaceTRUEAsian & -0.0375 & 0.0101 & -3.7155 & ** \\ 
  Income:EthRaceTRUEMulti & -0.0331 & 0.0192 & -1.7208 &   \\ 
  Income:EthRaceTRUENativeAm & -0.0163 & 0.0107 & -1.5276 &   \\ 
  Income:EthRaceTRUEOther & -0.0274 & 0.0080 & -3.4318 & ** \\ 
  Income:EthRaceTRUEPacific & -0.0240 & 0.0659 & -0.3647 &   \\ 
  Income:EthRaceTRUEWt & -0.0185 & 0.0033 & -5.5782 & *** \\ 
   \hline
\hline
\label{tab:SystemKWHEthnicity}
\end{longtable}
}
Table 
% \ref{tab:SystemKWHEthnicity}
shows similarities with the results from the orthodox electricity model.  Non-Hispanic Asians, and Hispanic Caucasians and Hispanic heads of households that identify as some other race all show lower electricity use per dollar of income.   This is consistent in sign with the parameter estimates in table 
% \ref{tab:ParOrthKWH}
, but the statistically significant deviations from non-Hispanic Caucasians for non Hispanic multi-ethnic heads of household has vanished.  The scale of the significant parameter estimates are also roughly the same magnitude but attenuated towards zero.

% latex table generated in R 3.0.1 by xtable 1.7-4 package
% Tue Apr  7 17:03:49 2015
{\small
\begin{longtable}{rrrrl}
\caption{System Estimation of kWh Model: Square Footage} \\ 
  \hline
 & Estimate & Std. Error & t value & Sig \\ 
  \hline
PSQFT:OwnSFDetached:HDD65 & 0.0001 & 0.0000 & 4.4109 & *** \\ 
  PSQFT:HDD65:OwnLgApartment & 0.0000 & 0.0001 & 0.8450 &   \\ 
  PSQFT:HDD65:OwnMobile & 0.0004 & 0.0001 & 4.6205 & *** \\ 
  PSQFT:HDD65:OwnSFAttached & 0.0000 & 0.0000 & 1.3603 &   \\ 
  PSQFT:HDD65:OwnSmApartment & 0.0001 & 0.0000 & 1.1709 &   \\ 
  PSQFT:HDD65:RentLgApartment & -0.0001 & 0.0000 & -1.3292 &   \\ 
  PSQFT:HDD65:RentMobile & 0.0005 & 0.0003 & 2.0238 &   \\ 
  PSQFT:HDD65:RentSFAttached & 0.0001 & 0.0001 & 1.0923 &   \\ 
  PSQFT:HDD65:RentSFDetached & 0.0001 & 0.0000 & 1.7233 &   \\ 
  PSQFT:HDD65:RentSmApartment & -0.0000 & 0.0000 & -0.2465 &   \\ 
  PSQFT:OwnSFDetached:CDD65 & 0.0007 & 0.0000 & 16.3870 & *** \\ 
  PSQFT:OwnLgApartment:CDD65 & 0.0000 & 0.0002 & 0.2207 &   \\ 
  PSQFT:OwnMobile:CDD65 & 0.0011 & 0.0002 & 6.7660 & *** \\ 
  PSQFT:OwnSFAttached:CDD65 & 0.0005 & 0.0001 & 5.3085 & *** \\ 
  PSQFT:OwnSmApartment:CDD65 & 0.0003 & 0.0002 & 1.8671 &   \\ 
  PSQFT:RentLgApartment:CDD65 & 0.0003 & 0.0001 & 2.5551 & . \\ 
  PSQFT:RentMobile:CDD65 & 0.0020 & 0.0009 & 2.2066 &   \\ 
  PSQFT:RentSFAttached:CDD65 & 0.0009 & 0.0002 & 4.9470 & *** \\ 
  PSQFT:RentSFDetached:CDD65 & 0.0008 & 0.0001 & 8.8123 & *** \\ 
  PSQFT:RentSmApartment:CDD65 & 0.0005 & 0.0001 & 3.5753 & ** \\ 
   \hline
\hline
\label{tab:SystemKWHSQFT}
\end{longtable}
}
Table 
% \ref{tab:SystemKWHSQFT}
shows the relationship between heating and cooling degree days per square foot on electricity use for each of the tenure and structure type combinations.  These results are particularly troublesome because heating and cooling degree days, in conjunction with square footage, are one of the most reliable explanatory variables for electricity use.  Space heating and cooling are two major end-uses in most households.  The parameter estimates show that these reliable explanatory variables are only significant in a few rare cases.  There is a strong relationship heating and cooling degree days for owned single family detached structures, but only rented mobile homes show statistically significant reactions to heating loads and only large apartments, owned or rented, for cooling loads.

This is very strong evidence that our square footage choice model needs improvement if it will be used as part of this electricity consumption system.
